---
title: "Super Bowl Commercials"
author: "Joe DiNoto"
date: "3/7/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, load packages.

```{r message=FALSE,warning=FALSE,error=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidytuesdayR)
library(scales)
```

Second, grab the data.

```{r message=FALSE,warning=FALSE,error=FALSE}
tuesdata <- tidytuesdayR::tt_load('2021-03-02')
youtube<- tuesdata$youtube
```

Look at the data.
```{r}
summary(youtube)
```
How many ads per year?
```{r}
youtube %>%
  count(year) %>%
  ggplot(aes(year,n))+
  geom_line() +
  scale_y_continuous(expand=c(0,0),limits = c(0,15))+
  labs(title="# of superbowl commercials per year",
       x="year",
       y="# of superbowl ads")
```
```{r}
youtube %>%
  filter(!is.na(brand))%>%
  mutate(brand=fct_lump(brand,4))%>%
  count(brand,year) %>%
  ggplot(aes(year,n)) +
  geom_col() +
  facet_grid(vars(brand))+
  labs(title="# of commercials per year, by category")
```
How well-liked (or disliked) are certain commercials?
```{r}
youtube%>%
  filter(dislike_count >0) %>%
  mutate(brand=fct_lump(brand,4))%>%
  mutate(L_D = like_count/dislike_count) %>%
  ggplot(aes(L_D,view_count,color=brand,size=view_count)) +
  geom_point()+
  scale_y_log10()+
  facet_grid(vars(use_sex))+
  labs(title="ratio of (like)/(dislike) vs. view count on youtube",
  subtitle = "FALSE = commercial does not use sex)")
```

It appears engagement is very low. The ratio of `likes/views` hovers just under `1%`, regardless of the brand, use of sex in the commercial, and/or brand.  
```{r}
youtube%>%
  filter(dislike_count >0) %>%
  mutate(brand=fct_lump(brand,4))%>%
  mutate(D_V = dislike_count/view_count) %>%
  ggplot(aes(D_V,view_count,color=brand,size=view_count)) +
  geom_point()+
  scale_y_log10()+
  scale_x_continuous(labels=percent)+
  facet_grid(vars(use_sex))+
  labs(title="ratio of (dislikes)/(views) vs. view count on youtube",
  subtitle = "FALSE = commercial does not use sex)")
```
```{r}
youtube%>%
  filter(dislike_count >0) %>%
  mutate(brand=fct_lump(brand,4))%>%
  mutate(L_V = like_count/view_count) %>%
  ggplot(aes(L_V,view_count,color=brand,size=view_count)) +
  geom_point()+
  scale_y_log10()+
  scale_x_continuous(labels=percent)+
  facet_grid(vars(use_sex))+
  labs(title="ratio of (likes)/(views) vs. view count on youtube",
  subtitle = "FALSE = commercial does not use sex)")
```
Following along with David Robinson
```{r}
youtube %>% 
  count(brand,sort=TRUE) %>%
  mutate(brand=fct_reorder(brand,n)) %>%
  ggplot(aes(n,brand))+
  geom_col()
```
```{r}
youtube %>%
  ggplot(aes(year,fill=brand))+
  geom_bar() +
  facet_wrap(~ brand)+
  theme(legend.position="none")
```

```{r}
youtube %>%
  gather(metric,value,contains("_count"))%>%
  ggplot(aes(value))+
  geom_histogram(binwidth=.5)+
  scale_x_log10(labels=comma)+
  facet_wrap(~metric)+
  labs(x="number of views")
```

```{r}
youtube %>%
  filter(!is.na(view_count))%>%
  mutate(brand=fct_reorder(brand,view_count))%>%
  ggplot(aes(view_count,brand, fill=funny))+
  geom_boxplot()+
  scale_x_log10(labels=comma)

```

median views per year
```{r}
youtube %>%
  filter(!is.na(view_count)) %>%
  group_by(year)%>%
  summarize(n=n(),
            median_views=median(view_count))%>%
  ggplot(aes(year,median_views))+
  geom_line()+
  geom_point(aes(size=n))+
  theme(legend.position = "none")+
  scale_y_log10(labels=comma)
```

```{r}
youtube %>%
  gather(category,value, funny:use_sex)%>%
  ggplot(aes(category,view_count, fill=value))+
  geom_boxplot()+
  scale_y_log10()
```

```{r}
youtube %>%
  filter(!is.na(view_count))%>%
  gather(category,value, funny:use_sex)%>%
  group_by(category,value) %>%
  summarize(n=n(),
            median_view_count=median(view_count)) %>%
  ggplot(aes(category,median_view_count, fill=value))+
  geom_col(position="dodge")
```

Correlations between value and the log of the view count.
```{r}
youtube %>%
  filter(!is.na(view_count))%>%
  gather(category,value, funny:use_sex)%>%
  group_by(category) %>%
  summarize(correlation = cor(value,log(view_count+1))) %>%
  arrange(desc(correlation))
```

```{r}
lm(log2(view_count) ~ danger+patriotic,
   data=youtube) %>%
  summary()
```

