---
title: "Volcano Analysis"
output: 
  html_document: 
    keep_md: yes
---

# Exploration and Visualization

Following along the example of David Robinson: 
https://github.com/dgrtwo/data-screencasts/blob/master/volcano-eruptions.Rmd

## Get the data

```{r message=FALSE, warning=FALSE, error=FALSE, cache=TRUE}
volcano <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv')
eruptions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/eruptions.csv')
events <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/events.csv')
tree_rings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/tree_rings.csv')
sulfur <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/sulfur.csv')
```

## Load some Libraries
```{r warning = FALSE, message=FALSE, error=FALSE}
library(tidyverse)
library(ggplot2)
```

## Changing the eruption year to a number

```{r}
volcano %>%
  mutate(last_eruption_year = as.numeric(last_eruption_year))
```

# Plot the number of eruptions by year as a histogram with log10 x axis.

```{r}
volcano %>%
  count(evidence_category, sort = TRUE)
```

```{r}
volcano %>%
  mutate(years_ago= 2020-as.numeric(last_eruption_year)) %>%
  ggplot(aes(years_ago+1, fill=evidence_category)) +
  geom_histogram()+
  scale_x_log10()
```

```{r}
volcano %>%
  count(region, sort = TRUE)
volcano %>%
  count(primary_volcano_type, sort = TRUE)
```
# Making a map using ggthemes

```{r}
library(ggthemes)
volcano %>%
  mutate(primary_volcano_type = str_remove(primary_volcano_type, "\\(.*\\)"), # remove the (s) from 
         # volcano types
         primary_volcano_type = fct_lump(primary_volcano_type, 6)) %>% # top 6 types + "other"
  ggplot(aes(longitude, latitude)) + # plot points based on long,lat coords
  borders() + # add country borders
  geom_point(aes(color = primary_volcano_type), size = 1, alpha=.4) + # color of dots = volcano type
  theme_map() + 
  labs(title = "Volcanoes of the World",
       color = "Type")
```

# more mapmaking

```{r}
library(leaflet)
library(glue)
library(htmlwidgets)
library(DT)
template <- "<>{ volcano_name }</p><p>{ primary_volcano_type }</p>"
volcano %>%
  mutate(transformed_pop = log2(population_within_5_km + 1),
         pop_color = colorNumeric(c("blue", "red"), transformed_pop)(transformed_pop)) %>%
  gather(key, value,
         volcano_name, primary_volcano_type, last_eruption_year,
         country,
         tectonic_settings,
         population_within_5_km) %>%
  mutate(key = str_to_title(str_replace_all(key, "_", " ")),
         key = paste0("<b>", key, "</b>")) %>%
  replace_na(list(value = "Unknown")) %>%
  nest(data = c(key, value)) %>%
  mutate(html = map(data,
                    knitr::kable,
                    format = "html",
                    escape = FALSE,
                    col.names = c("", ""))) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lat = ~ latitude,
                   lng = ~ longitude,
                   color = ~ pop_color,
                   popup = ~ html,
                   radius = 1) %>%
  addMeasure()
```

# Histogram of population within 5 km of any given volcano

```{r}
volcano %>%
  ggplot(aes(population_within_5_km)) +
  geom_histogram() +
  scale_x_log10()
```
```{r}
library(gganimate)
eruptions %>%
  ggplot(aes(2020 - start_year)) +
  geom_histogram() +
  scale_x_log10()
```
```{r}
eruptions %>%
  filter(start_year >= 1900,
         eruption_category != "Discredited Eruption") %>%
  mutate(years_ago = 2020-start_year) %>%
  ggplot(aes(years_ago)) +
  geom_histogram() +
  scale_x_log10()
```

# Machine Learning

Following along the example of Julia Silge:
https://juliasilge.com/blog/multinomial-volcano-eruptions/ 

## First, let's get the data.
```{r cache=TRUE, warning=FALSE, message=FALSE, error=FALSE}
volcano_raw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv")
```


## Count the number of primary volcano types
```{r}
volcano_raw %>%
  count(primary_volcano_type, sort = TRUE)
```

## Too many volcano types

See the code below to see how to reduce the number of volcano types to the two most common plus "other". 

```{r}
volcano_df <- volcano_raw %>% # create a new dataframe named volcano_df
  transmute( # transmute() is a dplyr functino that adds new variables 
    # and drops existing ones while preserving the number of rows. 
    # The New variables overwrite existing variables of the same name.
    volcano_type = case_when( # "stratovoclano" and "shield" are the new variable names 
      # to collect Stratovolcano(es) and "stratovolcano" into one variable. 
      #All other varialbes beyond that and "sheild" are simply named "other". 
      str_detect(primary_volcano_type, "Stratovolcano") ~ "Stratovolcano",
      str_detect(primary_volcano_type, "Shield") ~ "Shield",
      TRUE ~ "Other"
    ), # We still want to keep the following fields. 
    volcano_number, latitude, longitude, elevation,
    tectonic_settings, major_rock_1
  ) %>% 
  mutate_if(is.character, factor) #if a column contains characters, 
#it's automatically a factor. 
volcano_df %>%
  count(volcano_type, sort = TRUE)  #Now count the types of volcano according to the new grouping
```

## Let's build a map!

```{r warning=FALSE}
world <- map_data("world")

ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "white", fill = "gray50", size = 0.5, alpha = 0.2
  ) +
  geom_point(
    data = volcano_df,
    aes(longitude, latitude, color = volcano_type),
    alpha = 0.8
  ) +
  #theme_void(base_family = "IBMPlexSans") +
  labs(x = NULL, y = NULL, color = NULL)
```


# Bootstrap resample time!

```{r message=FALSE, warning=FALSE, error=FALSE}
library(tidymodels)
volcano_boot <- bootstraps(volcano_df)

volcano_boot
```


```{r message=FALSE, warning=FALSE, error=FALSE}
library(themis)

volcano_rec <- recipe(volcano_type ~ ., data = volcano_df) %>%
  update_role(volcano_number, new_role = "Id") %>%
  step_other(tectonic_settings) %>%
  step_other(major_rock_1) %>%
  step_dummy(tectonic_settings, major_rock_1) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors()) %>%
  step_smote(volcano_type)
```

```{r}
volcano_prep <- prep(volcano_rec)
juice(volcano_prep)
```

> *The ranger implementation for random forest can handle multinomial classification without any special handling.* - Julia Silge

```{r cache=TRUE}
rf_spec <- rand_forest(trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger")

volcano_wf <- workflow() %>%
  add_recipe(volcano_rec) %>%
  add_model(rf_spec)

volcano_wf
```

## Fitting the workflow to the resamples

```{r cache=TRUE}
volcano_res <- fit_resamples(
  volcano_wf,
  resamples = volcano_boot,
  control = control_resamples(save_pred = TRUE)
)
```

## Explore results

```{r}
volcano_res %>%
  collect_metrics()
```

## Confusion matrix

```{r}
volcano_res %>%
  collect_predictions() %>%
  conf_mat(volcano_type, .pred_class)
```

## Grouping by resample

```{r}
volcano_res %>%
  collect_predictions() %>%
  group_by(id) %>%
  ppv(volcano_type, .pred_class)
```

## Which variables are important? 

```{r message=FALSE, warning=FALSE, error=FALSE}
library(vip)

rf_spec %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(
    volcano_type ~ .,
    data = juice(volcano_prep) %>%
      select(-volcano_number) %>%
      janitor::clean_names()
  ) %>%
  vip(geom = "point")
```



```{r}
volcano_pred <- volcano_res %>%
  collect_predictions() %>%
  mutate(correct = volcano_type == .pred_class) %>%
  left_join(volcano_df %>%
    mutate(.row = row_number()))

volcano_pred
```

```{r}
ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
    color = "white", fill = "gray90", size = 0.05, alpha = 0.5
  ) +
  stat_summary_hex(
    data = volcano_pred,
    aes(longitude, latitude, z = as.integer(correct)),
    fun = "mean",
    alpha = 0.7, bins = 50
  ) +
  scale_fill_gradient(high = "cyan3", labels = scales::percent) +
  #theme_void(base_family = "IBMPlexSans") +
  labs(x = NULL, y = NULL, fill = "Percent classified\ncorrectly")
```

